---
title: "EcoAcc - Experiments"
author: "Kara Dobson"
date: "`r format(Sys.time(), '%d %B, %Y')`"
format: 
  html:
    code-fold: true
    code-line-numbers: true
    code-tools:
      source: false
    toc: true
embed-resources: true
dpi: 300
editor: visual
editor_options: 
  chunk_output_type: console
---

```{r, echo = F, message = F, warning = F}
# Load packages
library(tidyverse)
library(ggpubr)
library(deeptime)
library(patchwork)
library(plotly)
library(maps)
library(mapdata)
library(gridExtra)
library(viridis)

# Set path to turbo to get data
path_data = "/Volumes/seas-zhukai/proj-ecoacc/TRY_data/"
setwd(path_data)
# Load in data
traits <- read.csv(" exp_species_traits.csv")

# Set path to turbo to get data
path_data = "/Volumes/seas-zhukai/proj-ecoacc/TeRaCON/"
setwd(path_data)
# Load in data
CTI_sens_teracon <- read.csv(" CTI_sens_teracon.csv")
CTI_teracon <- read.csv(" CTI_teracon.csv")
tera <- read.csv(" teracon_clean.csv")
NPP_teracon <- read.csv(" eco_response_teracon.csv")
NPP_overall_teracon <- read.csv(" eco_response_overall_teracon.csv")
CTI_CPI_teracon <- read.csv(" CTI_CPI_teracon.csv")
niche_est_tera <- read.csv(" teracon_niche.csv")

# Set path to data
path_data = "/Volumes/seas-zhukai/proj-ecoacc/JRGCE/"
setwd(path_data)
# Load in data
CTI_sens_jrgce <- read.csv(" CTI_sens_jrgce.csv")
CTI_jrgce <- read.csv(" CTI_jrgce.csv")
jrgce <- read.csv(" jrgce_clean.csv")
jrgce <- jrgce %>%
  mutate(temp_treatment = if_else(str_detect(treatment, "T"), "warmed", "ambient"))
NPP_jrgce <- read.csv(" eco_response_jrgce.csv")
NPP_overall_jrgce <- read.csv(" eco_response_overall_jrgce.csv")
CTI_CPI_jrgce <- read.csv(" CTI_CPI_jrgce.csv")
niche_est_jrgce <- read.csv(" jrgce_niche.csv")

# Set path to data
path_data = "/Volumes/seas-zhukai/proj-ecoacc/PHACE/"
setwd(path_data)
# Load in data
CTI_sens_phace <- read.csv(" CTI_sens_phace.csv")
CTI_phace <- read.csv(" CTI_phace.csv")
phace <- read.csv(" phace_clean.csv")
NPP_phace <- read.csv(" eco_response_phace.csv")
NPP_overall_phace <- read.csv(" eco_response_overall_phace.csv")
CTI_CPI_phace <- read.csv(" CTI_CPI_phace.csv")
niche_est_phace <- read.csv(" phace_niche.csv")

# Set path to data
path_data = "/Volumes/seas-zhukai/proj-ecoacc/B4Warmed/"
setwd(path_data)
# Load in data
CTI_sens_b4 <- read.csv(" CTI_sens_b4warmed.csv")
CTI_b4 <- read.csv(" CTI_b4warmed.csv")
b4 <- read.csv(" b4warmed_clean.csv")
NPP_b4 <- read.csv(" eco_response_b4warmed.csv")
NPP_overall_b4 <- read.csv(" eco_response_overall_b4warmed.csv")
CTI_CPI_b4 <- read.csv(" CTI_CPI_b4warmed.csv")
niche_est_b4 <- read.csv(" b4warmed_niche.csv")

# Set path to data
path_data = "/Volumes/seas-zhukai/proj-ecoacc/OK/"
setwd(path_data)
# Load in data
CTI_sens_ok <- read.csv(" CTI_sens_ok.csv")
CTI_ok <- read.csv(" CTI_ok.csv")
ok <- read.csv(" ok_clean.csv")
NPP_ok <- read.csv(" eco_response_ok.csv")
NPP_overall_ok <- read.csv(" eco_response_overall_ok.csv")
CTI_CPI_ok <- read.csv(" CTI_CPI_ok.csv")
niche_est_ok <- read.csv(" ok_niche.csv")
```

Using community composition and ecosystem response data from various warming experiments, we can visualize changes in community structure (e.g., CTI) alongside changes in function (e.g., ANPP) due to warming.

## Map of the locations of each warming experiment

```{r, message = F, warning = F,fig.width = 10, fig.height = 6}
# Load data
us_map <- map_data("state")

# List of locations for each experiment
locations <- tibble(
  Experiment = c("TeRaCON", "B4Warmed CFC", "B4Warmed HWRC", "Oklahoma", "PHACE", "JRGCE"),
  lat = c(45, 46.7, 47.9, 35, 41.2, 37.4),
  long = c(-93, -92.5, -91.8, -97.5, -104.9, -122.2)
)

ggplot() +
  geom_polygon(data = us_map, aes(x = long, y = lat, group = group), fill = "lightgray", color = "black") +
  geom_point(data = locations, aes(x = long, y = lat, color = Experiment), size = 4) +
  coord_fixed(1.3) +
  scale_color_viridis(discrete=T) +
  theme_minimal() +
  theme(legend.text = element_text(size=12),
        legend.title = element_text(size=12),
        axis.text = element_text(size=12),
        axis.title = element_text(size=12))
```

## CTI and NPP

### Changes in CTI over time between warmed and ambient treatments from 6 in situ warming experiments

```{r,message = F, warning = F, fig.width = 10, fig.height = 6}
# CTI sensitivity figures
CTI_tera_plot <- ggplot(CTI_sens_teracon, aes(x = year, y = sensitivity)) +
  geom_smooth() +
  labs(x = "Year", y = "TeRaCON") +
  scale_x_continuous(breaks = seq(2012, 2023, by = 2)) +
  theme_bw()
CTI_jrgce_plot <- ggplot(CTI_sens_jrgce, aes(x = year, y = sensitivity)) +
  geom_smooth() +
  labs(x = "Year", y = "JRGCE", title="CTI (Warmed - Ambient)") +
  scale_x_continuous(breaks = seq(1998, 2014, by = 3)) +
  theme_bw()
CTI_phace_plot <- ggplot(CTI_sens_phace, aes(x = year, y = sensitivity)) +
  geom_smooth() +
  labs(x = "Year", y = "PHACE") +
  scale_x_continuous(breaks = seq(2007, 2013)) +
  theme_bw()
CTI_sens_b4_cfc <- CTI_sens_b4 %>%
  filter(site == "CFC")
CTI_sens_b4_hwrc <- CTI_sens_b4 %>%
  filter(site == "HWRC")
CTI_b4_cfc_plot <- ggplot(CTI_sens_b4_cfc, aes(x = year, y = sensitivity_high_temp)) +
  geom_smooth() +
  labs(x = "Year", y = "B4Warmed CFC") +
  scale_x_continuous(breaks = seq(2008, 2021, by = 2)) +
  theme_bw()
CTI_b4_hwrc_plot <- ggplot(CTI_sens_b4_hwrc, aes(x = year, y = sensitivity_high_temp)) +
  geom_smooth() +
  labs(x = "Year", y = "B4Warmed HWRC") +
  scale_x_continuous(breaks = seq(2008, 2021, by = 2)) +
  theme_bw()
CTI_ok_plot <- ggplot(CTI_sens_ok, aes(x = year, y = sensitivity)) +
  geom_smooth() +
  labs(x = "Year", y = "Oklahoma") +
  #scale_x_continuous(breaks = seq(2007, 2013)) +
  theme_bw()

# CTI figures
CTI_tera_plot2 <- ggplot(CTI_teracon, aes(x = year, y = CTI, color = temp_treatment)) +
  geom_smooth() +
  labs(x = "Year", y = "CTI", title="TeRaCON") +
  scale_x_continuous(breaks = seq(2012, 2023, by = 2)) +
  scale_color_manual(name = "Treatment",
                     labels = c("Ambient","Warmed"),
                     values = c("blue","red")) +
  theme_bw()
CTI_jrgce_plot2 <- ggplot(CTI_jrgce, aes(x = year, y = CTI, color = temp_treatment)) +
  geom_smooth() +
  labs(x = "Year", y = "CTI",title="JRGCE") +
  scale_x_continuous(breaks = seq(1998, 2014, by = 3)) +
  scale_color_manual(name = "Treatment",
                     labels = c("Ambient","Warmed"),
                     values = c("blue","red")) +
  theme_bw()
CTI_phace_plot2 <- ggplot(CTI_phace, aes(x = year, y = CTI, color = temp_treatment)) +
  geom_smooth() +
  labs(x = "Year", y = "CTI",title="PHACE") +
  scale_x_continuous(breaks = seq(2007, 2013)) +
  scale_color_manual(name = "Treatment",
                     labels = c("Ambient","Warmed"),
                     values = c("blue","red")) +
  theme_bw()
CTI_b4_cfc <- CTI_b4 %>%
  filter(site == "CFC") %>%
  filter(temp_treatment == "amb" | temp_treatment == 3.4)
CTI_b4_hwrc <- CTI_b4 %>%
  filter(site == "HWRC") %>%
  filter(temp_treatment == "amb" | temp_treatment == 3.4)
CTI_b4_cfc_plot2 <- ggplot(CTI_b4_cfc, aes(x = year, y = CTI, color = temp_treatment)) +
  geom_smooth() +
  labs(x = "Year", y = "CTI",title="B4Warmed CFC") +
  scale_x_continuous(breaks = seq(2008, 2021, by = 2)) +
  scale_color_manual(name = "Treatment",
                     breaks=c("amb","3.4"),
                     labels = c("Ambient","Warmed"),
                     values = c("blue","red")) +
  theme_bw()
CTI_b4_hwrc_plot2 <- ggplot(CTI_b4_hwrc, aes(x = year, y = CTI, color = temp_treatment)) +
  geom_smooth() +
  labs(x = "Year", y = "CTI",title="B4Warmed HWRC") +
  scale_x_continuous(breaks = seq(2008, 2021, by = 2)) +
  scale_color_manual(name = "Treatment",
                     breaks=c("amb","3.4"),
                     labels = c("Ambient","Warmed"),
                     values = c("blue","red")) +
  theme_bw()
CTI_ok_plot2 <- ggplot(CTI_ok, aes(x = year, y = CTI, color = temp_treatment)) +
  geom_smooth() +
  labs(x = "Year", y = "CTI",title="Oklahoma") +
  #scale_x_continuous(breaks = seq(2007, 2013)) +
  scale_color_manual(name = "Treatment",
                     labels = c("Ambient","Warmed"),
                     values = c("blue","red")) +
  theme_bw()

# Biomass figures
NPP_teracon <- NPP_teracon %>%
  filter(variable == "mean_ab_bio")
npp_tera_plot <- ggplot(NPP_teracon, aes(x = year, y = sensitivity)) +
  geom_smooth() +
  labs(x = "Year", y = NULL) +
  scale_x_continuous(breaks = seq(2012, 2023, by = 2)) +
  theme_bw()
npp_jrgce_plot <- ggplot(NPP_jrgce, aes(x = year, y = sensitivity)) +
  geom_smooth() +
  labs(x = "Year", y = NULL, title="Biomass\n(Warmed - Ambient)") +
  scale_x_continuous(breaks = seq(1998, 2014, by = 3)) +
  theme_bw()
npp_phace_plot <- ggplot(NPP_phace, aes(x = year, y = sensitivity)) +
  geom_smooth() +
  labs(x = "Year", y = NULL) +
  scale_x_continuous(breaks = seq(2007, 2013)) +
  theme_bw()
NPP_b4_cfc <- NPP_b4 %>%
  filter(site == "CFC")
NPP_b4_hwrc <- NPP_b4 %>%
  filter(site == "HWRC")
npp_b4_cfc_plot <- ggplot(NPP_b4_cfc, aes(x = year, y = sensitivity_high_temp)) +
  geom_smooth() +
  labs(x = "Year", y = NULL) +
  scale_x_continuous(breaks = seq(2008, 2021, by = 2)) +
  theme_bw()
npp_b4_hwrc_plot <- ggplot(NPP_b4_hwrc, aes(x = year, y = sensitivity_high_temp)) +
  geom_smooth() +
  labs(x = "Year", y = NULL) +
  scale_x_continuous(breaks = seq(2008, 2021, by = 2)) +
  theme_bw()
npp_ok_plot <- ggplot(NPP_ok, aes(x = year, y = sensitivity)) +
  geom_smooth() +
  labs(x = "Year", y = NULL) +
  #scale_x_continuous(breaks = seq(2007, 2013)) +
  theme_bw()

# Ambient temp change figures
temp_tera_plot <- ggplot(CTI_sens_teracon, aes(x = year, y = mean_C_temp_summer)) +
  geom_smooth() +
  labs(x = "Year", y = NULL) +
  scale_x_continuous(breaks = seq(2012, 2023, by = 2)) +
  theme_bw()
temp_jrgce_plot <- ggplot(CTI_sens_jrgce, aes(x = year, y = mean_C_temp_summer)) +
  geom_smooth() +
  labs(x = "Year", title = "Ambient temperature (°C)",y=NULL) +
  scale_x_continuous(breaks = seq(1998, 2014, by = 3)) +
  theme_bw()

combined <- wrap_plots(CTI_jrgce_plot2,CTI_phace_plot2,CTI_tera_plot2,
           CTI_ok_plot2,CTI_b4_cfc_plot2,CTI_b4_hwrc_plot2,
           ncol = 3) & theme(legend.position = "right")
combined + plot_layout(guides = "collect")
```

### The effect of warming on CTI and NPP over time, including ambient temperature changes (when it was available in the experimental data)

```{r,message = F, warning = F, fig.width = 10, fig.height = 9}
ggarrange2(CTI_jrgce_plot,npp_jrgce_plot,temp_jrgce_plot,
           CTI_phace_plot,npp_phace_plot,
           CTI_tera_plot,npp_tera_plot,temp_tera_plot,
           CTI_ok_plot,npp_ok_plot,
           CTI_b4_cfc_plot,npp_b4_cfc_plot,
           CTI_b4_hwrc_plot,npp_b4_hwrc_plot,
           layout = matrix(c(1, 2, 3,
                            4, 5, 0,
                            6, 7, 8,
                            9, 10, 0,
                            11, 12, 0,
                            13, 14, 0), nrow = 6, byrow = TRUE)
)
```

### Relationship between CTI and aboveground biomass production each year

```{r, message = F, warning = F, fig.width = 14, fig.height = 6}
##### Fig: correlating biomass w/ CTI
# TeRaCON
CTI_yearly_avg_tera <- CTI_teracon %>%
  group_by(year,temp_treatment) %>%
  summarize(mean_CTI = mean(CTI))
NPP_CTI_teracon <- left_join(CTI_yearly_avg_tera, NPP_overall_teracon, by=c("year","temp_treatment"))
tera_scatter <- ggscatter(NPP_CTI_teracon, x = "mean_CTI", y = "mean_ab_bio", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "CTI", ylab = "Biomass", title="TeRaCON")
# JRGCE
CTI_yearly_avg_jrgce <- CTI_jrgce %>%
  group_by(year,temp_treatment) %>%
  summarize(mean_CTI = mean(CTI))
NPP_CTI_jrgce <- left_join(CTI_yearly_avg_jrgce, NPP_overall_jrgce, by=c("year","temp_treatment"))
jrgce_scatter <- ggscatter(NPP_CTI_jrgce, x = "mean_CTI", y = "mean_ab_bio", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "CTI", ylab = "Biomass",title="JRGCE")
# PHACE
CTI_yearly_avg_phace <- CTI_phace %>%
  group_by(year,temp_treatment) %>%
  summarize(mean_CTI = mean(CTI))
NPP_CTI_phace <- left_join(CTI_yearly_avg_phace, NPP_overall_phace, by=c("year","temp_treatment"))
phace_scatter <- ggscatter(NPP_CTI_phace, x = "mean_CTI", y = "mean_ab_bio", 
                           add = "reg.line", conf.int = TRUE, 
                           cor.coef = TRUE, cor.method = "pearson",
                           xlab = "CTI", ylab = "Biomass",title="PHACE")
# B4Warmed
CTI_yearly_avg_b4_cfc <- CTI_b4 %>%
  filter(site == "CFC") %>%
  group_by(year,temp_treatment) %>%
  summarize(mean_CTI = mean(CTI))
CTI_yearly_avg_b4_hwrc <- CTI_b4 %>%
  filter(site == "HWRC") %>%
  group_by(year,temp_treatment) %>%
  summarize(mean_CTI = mean(CTI))
NPP_overall_b4_cfc <- NPP_overall_b4 %>%
  filter(site == "CFC")
NPP_overall_b4_hwrc <- NPP_overall_b4 %>%
  filter(site == "HWRC")
NPP_CTI_b4_cfc <- left_join(CTI_yearly_avg_b4_cfc, NPP_overall_b4_cfc, by=c("year","temp_treatment"))
NPP_CTI_b4_hwrc <- left_join(CTI_yearly_avg_b4_hwrc, NPP_overall_b4_hwrc, by=c("year","temp_treatment"))
b4_scatter_cfc <- ggscatter(NPP_CTI_b4_cfc, x = "mean_CTI", y = "mean_ab_bio", 
                           add = "reg.line", conf.int = TRUE, 
                           cor.coef = TRUE, cor.method = "pearson",
                           xlab = "CTI", ylab = "Biomass",title="B4Warmed CFC")
b4_scatter_hwrc <- ggscatter(NPP_CTI_b4_hwrc, x = "mean_CTI", y = "mean_ab_bio", 
                           add = "reg.line", conf.int = TRUE, 
                           cor.coef = TRUE, cor.method = "pearson",
                           xlab = "CTI", ylab = "Biomass",title="B4Warmed HWRC")
# OK
CTI_yearly_avg_ok <- CTI_ok %>%
  group_by(year,temp_treatment) %>%
  summarize(mean_CTI = mean(CTI))
NPP_CTI_ok <- left_join(CTI_yearly_avg_ok, NPP_overall_ok, by=c("year","temp_treatment"))
ok_scatter <- ggscatter(NPP_CTI_ok, x = "mean_CTI", y = "mean_ab_bio", 
                           add = "reg.line", conf.int = TRUE, 
                           cor.coef = TRUE, cor.method = "pearson",
                           xlab = "CTI", ylab = "Biomass",title="Oklahoma")

# plot
ggarrange(jrgce_scatter,phace_scatter,tera_scatter,ok_scatter,b4_scatter_cfc,b4_scatter_hwrc,
          ncol=3,nrow=2)
```

## Treatment arrows

### Each arrow represents one year. The arrow points in the direction of the treatment effect (ambient –\> warmed) and is colored by the year (yellow = most recent)

```{r, warning=F,message=F, fig.width = 14, fig.height = 7}
##### Fig: Arrows pointing from ambient to warmed for each year #####
arrow_teracon <- ggplot(CTI_CPI_teracon) +
  geom_segment(aes(x = CTI_HTamb, y = CPI_HTamb, 
                   xend = CTI_HTelv, yend = CPI_HTelv,
                   color = year),
               arrow = arrow(length = unit(0.1, "inches"))) +
  geom_point(aes(x = CTI_HTamb, y = CPI_HTamb), color = "black") +
  geom_point(aes(x = CTI_HTelv, y = CPI_HTelv), color = "red") +
  labs(x = "CTI", y = "CPI", title = "TeRaCON") +
  scale_color_viridis_c(option = "viridis") +
  theme_minimal()

arrow_jrgce <- ggplot(CTI_CPI_jrgce) +
  geom_segment(aes(x = CTI_ambient, y = CPI_ambient, 
                   xend = CTI_warmed, yend = CPI_warmed,
                   color = year),
               arrow = arrow(length = unit(0.1, "inches"))) +
  geom_point(aes(x = CTI_ambient, y = CPI_ambient), color = "black") +
  geom_point(aes(x = CTI_warmed, y = CPI_warmed), color = "red") +
  labs(x = "CTI", y = "CPI", title = "JRGCE") +
  scale_color_viridis_c(option = "viridis") +
  theme_minimal()

arrow_phace <- ggplot(CTI_CPI_phace) +
  geom_segment(aes(x = CTI_ambient, y = CPI_ambient, 
                   xend = CTI_warmed, yend = CPI_warmed,
                   color = year),
               arrow = arrow(length = unit(0.1, "inches"))) +
  geom_point(aes(x = CTI_ambient, y = CPI_ambient), color = "black") +
  geom_point(aes(x = CTI_warmed, y = CPI_warmed), color = "red") +
  labs(x = "CTI", y = "CPI", title = "PHACE") +
  scale_color_viridis_c(option = "viridis") +
  theme_minimal()

CTI_CPI_b4_cfc <- CTI_CPI_b4 %>%
  filter(site == "CFC")
CTI_CPI_b4_hwrc <- CTI_CPI_b4 %>%
  filter(site == "HWRC")
arrow_b4_cfc <- ggplot(CTI_CPI_b4_cfc) +
  geom_segment(aes(x = CTI_amb, y = CPI_amb, 
                   xend = CTI_3.4, yend = CPI_3.4,
                   color = year),
               arrow = arrow(length = unit(0.1, "inches"))) +
  geom_point(aes(x = CTI_amb, y = CPI_amb), color = "black") +
  geom_point(aes(x = CTI_3.4, y = CPI_3.4), color = "red") +
  labs(x = "CTI", y = "CPI", title = "B4Warmed CFC") +
  scale_color_viridis_c(option = "viridis") +
  theme_minimal()
arrow_b4_hwrc <- ggplot(CTI_CPI_b4_hwrc) +
  geom_segment(aes(x = CTI_amb, y = CPI_amb, 
                   xend = CTI_3.4, yend = CPI_3.4,
                   color = year),
               arrow = arrow(length = unit(0.1, "inches"))) +
  geom_point(aes(x = CTI_amb, y = CPI_amb), color = "black") +
  geom_point(aes(x = CTI_3.4, y = CPI_3.4), color = "red") +
  labs(x = "CTI", y = "CPI", title = "B4Warmed HWRC") +
  scale_color_viridis_c(option = "viridis") +
  theme_minimal()

arrow_ok <- ggplot(CTI_CPI_ok) +
  geom_segment(aes(x = CTI_ambient, y = CPI_ambient, 
                   xend = CTI_warmed, yend = CPI_warmed,
                   color = year),
               arrow = arrow(length = unit(0.1, "inches"))) +
  geom_point(aes(x = CTI_ambient, y = CPI_ambient), color = "black") +
  geom_point(aes(x = CTI_warmed, y = CPI_warmed), color = "red") +
  labs(x = "CTI", y = "CPI", title = "Oklahoma") +
  scale_color_viridis_c(option = "viridis") +
  theme_minimal()

ggarrange(arrow_jrgce,arrow_phace,arrow_teracon,arrow_ok,arrow_b4_cfc,arrow_b4_hwrc,
          ncol = 3, nrow=2)
```

### Arrows overlaid on all species climate niches for each experiment

```{r, warning=F,message=F,fig.width = 12, fig.height = 7}
##### Fig: Species climate niche overlayed w/ treatment effects
# TeRaCON
niche_est_tera_avg <- niche_est_tera[,c(1,6,7)]
niche_est_tera_avg <- niche_est_tera_avg %>%
  distinct()
teracon_niche <- ggplot(CTI_CPI_teracon) +
  geom_segment(aes(x = CTI_HTamb, y = CPI_HTamb, 
                   xend = CTI_HTelv, yend = CPI_HTelv,
                   color = year),
               arrow = arrow(length = unit(0.1, "inches"))) +
  geom_point(aes(x = CTI_HTamb, y = CPI_HTamb), color = "black") +
  geom_point(aes(x = CTI_HTelv, y = CPI_HTelv), color = "red") +
  geom_point(data=niche_est_tera_avg, aes(x = temp_niche, y = precip_niche)) +
  labs(x = "Temperature", y = "Precipitation", title = "TeRaCON") +
  scale_color_viridis_c(option = "viridis") +
  theme_minimal()

# JRGCE
niche_est_jrgce_avg <- niche_est_jrgce[,c(1,6,7)]
niche_est_jrgce_avg <- niche_est_jrgce_avg %>%
  distinct()
jrgce_niche <- ggplot(CTI_CPI_jrgce) +
  geom_segment(aes(x = CTI_ambient, y = CPI_ambient, 
                   xend = CTI_warmed, yend = CPI_warmed,
                   color = year),
               arrow = arrow(length = unit(0.1, "inches"))) +
  geom_point(aes(x = CTI_ambient, y = CPI_ambient), color = "black") +
  geom_point(aes(x = CTI_warmed, y = CPI_warmed), color = "red") +
  geom_point(data=niche_est_jrgce_avg, aes(x = temp_niche, y = precip_niche)) +
  labs(x = "Temperature", y = "Precipitation", title = "JRGCE") +
  scale_color_viridis_c(option = "viridis") +
  theme_minimal()

# PHACE
niche_est_phace_avg <- niche_est_phace[,c(1,6,7)]
niche_est_phace_avg <- niche_est_phace_avg %>%
  distinct()
phace_niche <- ggplot(CTI_CPI_phace) +
  geom_segment(aes(x = CTI_ambient, y = CPI_ambient, 
                   xend = CTI_warmed, yend = CPI_warmed,
                   color = year),
               arrow = arrow(length = unit(0.1, "inches"))) +
  geom_point(aes(x = CTI_ambient, y = CPI_ambient), color = "black") +
  geom_point(aes(x = CTI_warmed, y = CPI_warmed), color = "red") +
  geom_point(data=niche_est_phace_avg, aes(x = temp_niche, y = precip_niche)) +
  labs(x = "Temperature", y = "Precipitation", title = "PHACE") +
  scale_color_viridis_c(option = "viridis") +
  theme_minimal()

# B4Warmed
niche_est_b4_cfc <- niche_est_b4 %>%
  filter(site == "CFC")
CTI_CPI_b4_cfc <- CTI_CPI_b4 %>%
  filter(site == "CFC")
niche_est_b4_hwrc <- niche_est_b4 %>%
  filter(site == "HWRC")
CTI_CPI_b4_hwrc <- CTI_CPI_b4 %>%
  filter(site == "HWRC")
niche_est_b4_avg_cfc <- niche_est_b4_cfc[,c(1,7,8)]
niche_est_b4_avg_cfc <- niche_est_b4_avg_cfc %>%
  distinct()
niche_est_b4_avg_hwrc <- niche_est_b4_hwrc[,c(1,7,8)]
niche_est_b4_avg_hwrc <- niche_est_b4_avg_hwrc %>%
  distinct()
b4_niche_cfc <- ggplot(CTI_CPI_b4_cfc) +
  geom_segment(aes(x = CTI_amb, y = CPI_amb, 
                   xend = CTI_3.4, yend = CPI_3.4,
                   color = year),
               arrow = arrow(length = unit(0.1, "inches"))) +
  geom_point(aes(x = CTI_amb, y = CPI_amb), color = "black") +
  geom_point(aes(x = CTI_3.4, y = CPI_3.4), color = "red") +
  geom_point(data=niche_est_b4_avg_cfc, aes(x = temp_niche, y = precip_niche)) +
  labs(x = "Temperature", y = "Precipitation", title = "B4Warmed CFC") +
  scale_color_viridis_c(option = "viridis") +
  theme_minimal()
b4_niche_hwrc <- ggplot(CTI_CPI_b4_hwrc) +
  geom_segment(aes(x = CTI_amb, y = CPI_amb, 
                   xend = CTI_3.4, yend = CPI_3.4,
                   color = year),
               arrow = arrow(length = unit(0.1, "inches"))) +
  geom_point(aes(x = CTI_amb, y = CPI_amb), color = "black") +
  geom_point(aes(x = CTI_3.4, y = CPI_3.4), color = "red") +
  geom_point(data=niche_est_b4_avg_hwrc, aes(x = temp_niche, y = precip_niche)) +
  labs(x = "Temperature", y = "Precipitation", title = "B4Warmed HWRC") +
  scale_color_viridis_c(option = "viridis") +
  theme_minimal()

# PHACE
niche_est_ok_avg <- niche_est_ok[,c(1,6,7)]
niche_est_ok_avg <- niche_est_ok_avg %>%
  distinct()
ok_niche <- ggplot(CTI_CPI_ok) +
  geom_segment(aes(x = CTI_ambient, y = CPI_ambient, 
                   xend = CTI_warmed, yend = CPI_warmed,
                   color = year),
               arrow = arrow(length = unit(0.1, "inches"))) +
  geom_point(aes(x = CTI_ambient, y = CPI_ambient), color = "black") +
  geom_point(aes(x = CTI_warmed, y = CPI_warmed), color = "red") +
  geom_point(data=niche_est_ok_avg, aes(x = temp_niche, y = precip_niche)) +
  labs(x = "Temperature", y = "Precipitation", title = "Oklahoma") +
  scale_color_viridis_c(option = "viridis") +
  theme_minimal()

ggarrange(jrgce_niche,phace_niche,teracon_niche,ok_niche,b4_niche_cfc,b4_niche_hwrc,
          ncol=3,nrow=2)
```

### Centering the ambient treatment at 0 to better show directional shifts over time

```{r, warning=F,message=F,fig.width = 12, fig.height = 7}
##### Fig: Adjusted arrow figure for year and treatment effects
CTI_CPI_tera_sens <- CTI_CPI_teracon %>%
  mutate(CTI_sens = CTI_HTelv - CTI_HTamb,
         CPI_sens = CPI_HTelv - CPI_HTamb)
CTI_CPI_tera_sens$CTI_ambient_scaled <- 0
CTI_CPI_tera_sens$CPI_ambient_scaled <- 0

CTI_CPI_jrgce_sens <- CTI_CPI_jrgce %>%
  mutate(CTI_sens = CTI_warmed - CTI_ambient,
         CPI_sens = CPI_warmed - CPI_ambient)
CTI_CPI_jrgce_sens$CTI_ambient_scaled <- 0
CTI_CPI_jrgce_sens$CPI_ambient_scaled <- 0

CTI_CPI_phace_sens <- CTI_CPI_phace %>%
  mutate(CTI_sens = CTI_warmed - CTI_ambient,
         CPI_sens = CPI_warmed - CPI_ambient)
CTI_CPI_phace_sens$CTI_ambient_scaled <- 0
CTI_CPI_phace_sens$CPI_ambient_scaled <- 0

CTI_CPI_b4_sens_cfc <- CTI_CPI_b4 %>%
  filter(site == "CFC") %>%
  mutate(CTI_sens = CTI_3.4 - CTI_amb,
         CPI_sens = CPI_3.4 - CPI_amb)
CTI_CPI_b4_sens_cfc$CTI_ambient_scaled <- 0
CTI_CPI_b4_sens_cfc$CPI_ambient_scaled <- 0

CTI_CPI_b4_sens_hwrc <- CTI_CPI_b4 %>%
  filter(site == "HWRC") %>%
  mutate(CTI_sens = CTI_3.4 - CTI_amb,
         CPI_sens = CPI_3.4 - CPI_amb)
CTI_CPI_b4_sens_hwrc$CTI_ambient_scaled <- 0
CTI_CPI_b4_sens_hwrc$CPI_ambient_scaled <- 0

CTI_CPI_ok_sens <- CTI_CPI_ok %>%
  mutate(CTI_sens = CTI_warmed - CTI_ambient,
         CPI_sens = CPI_warmed - CPI_ambient)
CTI_CPI_ok_sens$CTI_ambient_scaled <- 0
CTI_CPI_ok_sens$CPI_ambient_scaled <- 0

arrow_sens_tera <- ggplot(CTI_CPI_tera_sens) +
  geom_segment(aes(x = CTI_ambient_scaled, y = CPI_ambient_scaled, 
                   xend = CTI_sens, yend = CPI_sens,
                   color = year),
               arrow = arrow(length = unit(0.1, "inches"))) +
  geom_point(aes(x = CTI_ambient_scaled, y = CPI_ambient_scaled), color = "black") +
  geom_point(aes(x = CTI_sens, y = CPI_sens), color = "red") +
  labs(x = "CTI", y = "CPI", title = "TeRaCON") +
  scale_color_viridis_c(option = "viridis") +
  #ylim(-50,40) +
  #xlim(-0.35,0.45) +
  theme_minimal()

arrow_sens_jrgce <- ggplot(CTI_CPI_jrgce_sens) +
  geom_segment(aes(x = CTI_ambient_scaled, y = CPI_ambient_scaled, 
                   xend = CTI_sens, yend = CPI_sens,
                   color = year),
               arrow = arrow(length = unit(0.1, "inches"))) +
  geom_point(aes(x = CTI_ambient_scaled, y = CPI_ambient_scaled), color = "black") +
  geom_point(aes(x = CTI_sens, y = CPI_sens), color = "red") +
  labs(x = "CTI", y = "CPI", title = "JRGCE") +
  scale_color_viridis_c(option = "viridis") +
  #ylim(-50,40) +
  #xlim(-0.35,0.45) +
  theme_minimal()

arrow_sens_phace <- ggplot(CTI_CPI_phace_sens) +
  geom_segment(aes(x = CTI_ambient_scaled, y = CPI_ambient_scaled, 
                   xend = CTI_sens, yend = CPI_sens,
                   color = year),
               arrow = arrow(length = unit(0.1, "inches"))) +
  geom_point(aes(x = CTI_ambient_scaled, y = CPI_ambient_scaled), color = "black") +
  geom_point(aes(x = CTI_sens, y = CPI_sens), color = "red") +
  labs(x = "CTI", y = "CPI", title = "PHACE") +
  scale_color_viridis_c(option = "viridis") +
  #ylim(-50,40) +
  #xlim(-0.35,0.45) +
  theme_minimal()

arrow_sens_b4_cfc <- ggplot(CTI_CPI_b4_sens_cfc) +
  geom_segment(aes(x = CTI_ambient_scaled, y = CPI_ambient_scaled, 
                   xend = CTI_sens, yend = CPI_sens,
                   color = year),
               arrow = arrow(length = unit(0.1, "inches"))) +
  geom_point(aes(x = CTI_ambient_scaled, y = CPI_ambient_scaled), color = "black") +
  geom_point(aes(x = CTI_sens, y = CPI_sens), color = "red") +
  labs(x = "CTI", y = "CPI", title = "B4Warmed CFC") +
  scale_color_viridis_c(option = "viridis") +
  #ylim(-50,40) +
  #xlim(-0.35,0.45) +
  theme_minimal()

arrow_sens_b4_hwrc <- ggplot(CTI_CPI_b4_sens_hwrc) +
  geom_segment(aes(x = CTI_ambient_scaled, y = CPI_ambient_scaled, 
                   xend = CTI_sens, yend = CPI_sens,
                   color = year),
               arrow = arrow(length = unit(0.1, "inches"))) +
  geom_point(aes(x = CTI_ambient_scaled, y = CPI_ambient_scaled), color = "black") +
  geom_point(aes(x = CTI_sens, y = CPI_sens), color = "red") +
  labs(x = "CTI", y = "CPI", title = "B4Warmed HWRC") +
  scale_color_viridis_c(option = "viridis") +
  #ylim(-50,40) +
  #xlim(-0.35,0.45) +
  theme_minimal()

arrow_sens_ok <- ggplot(CTI_CPI_ok_sens) +
  geom_segment(aes(x = CTI_ambient_scaled, y = CPI_ambient_scaled, 
                   xend = CTI_sens, yend = CPI_sens,
                   color = year),
               arrow = arrow(length = unit(0.1, "inches"))) +
  geom_point(aes(x = CTI_ambient_scaled, y = CPI_ambient_scaled), color = "black") +
  geom_point(aes(x = CTI_sens, y = CPI_sens), color = "red") +
  labs(x = "CTI", y = "CPI", title = "Oklahoma") +
  scale_color_viridis_c(option = "viridis") +
  #ylim(-50,40) +
  #xlim(-0.35,0.45) +
  theme_minimal()

ggarrange(arrow_sens_jrgce,arrow_sens_phace,arrow_sens_tera,arrow_sens_ok,arrow_sens_b4_cfc,arrow_sens_b4_hwrc,
          ncol = 3, nrow=2)
```

### Treatment effect for the final year of all experiments

```{r, warning=F,message=F,fig.width = 8, fig.height = 5}
# Filter to maximum year per experiment
tera_max <- CTI_CPI_tera_sens %>%
  filter(year == max(year))
jrgce_max <- CTI_CPI_jrgce_sens %>%
  filter(year == max(year))
phace_max <- CTI_CPI_phace_sens %>%
  filter(year == max(year))
ok_max <- CTI_CPI_ok_sens %>%
  filter(year == max(year))
b4_cfc_max <- CTI_CPI_b4_sens_cfc %>%
  filter(year == max(year))
b4_hwrc_max <- CTI_CPI_b4_sens_hwrc %>%
  filter(year == max(year))

cols <- c("jrgce_col"="#332288","phace_col"="#117733","tera_col"="#882255",
         "ok_col" =  "#88CCEE", "b4_cfc_col" = "#AA4499", "b4_hwrc_col" = "#DDCC77")
ggplot() +
  geom_segment(data=tera_max,
               aes(x = CTI_ambient_scaled, y = CPI_ambient_scaled,
                   xend = CTI_sens, yend = CPI_sens, color = "tera_col"),
               arrow = arrow(length = unit(0.1, "inches"))) +
  geom_point(data=tera_max,
             aes(x = CTI_ambient_scaled, y = CPI_ambient_scaled), color = "black") +
  geom_point(data=tera_max,
             aes(x = CTI_sens, y = CPI_sens), color = "red") +
  geom_segment(data=phace_max,
               aes(x = CTI_ambient_scaled, y = CPI_ambient_scaled,
                   xend = CTI_sens, yend = CPI_sens, color = "phace_col"),
               arrow = arrow(length = unit(0.1, "inches"))) +
  geom_point(data=phace_max,
             aes(x = CTI_ambient_scaled, y = CPI_ambient_scaled), color = "black") +
  geom_point(data=phace_max,
             aes(x = CTI_sens, y = CPI_sens), color = "red") +
  geom_segment(data=jrgce_max,
               aes(x = CTI_ambient_scaled, y = CPI_ambient_scaled,
                   xend = CTI_sens, yend = CPI_sens, color = "jrgce_col"),
               arrow = arrow(length = unit(0.1, "inches"))) +
  geom_point(data=jrgce_max,
             aes(x = CTI_ambient_scaled, y = CPI_ambient_scaled), color = "black") +
  geom_point(data=jrgce_max,
             aes(x = CTI_sens, y = CPI_sens), color = "red") +
   geom_segment(data=ok_max,
               aes(x = CTI_ambient_scaled, y = CPI_ambient_scaled,
                   xend = CTI_sens, yend = CPI_sens, color = "ok_col"),
               arrow = arrow(length = unit(0.1, "inches"))) +
  geom_point(data=ok_max,
             aes(x = CTI_ambient_scaled, y = CPI_ambient_scaled), color = "black") +
  geom_point(data=ok_max,
             aes(x = CTI_sens, y = CPI_sens), color = "red") +
  geom_segment(data=b4_cfc_max,
               aes(x = CTI_ambient_scaled, y = CPI_ambient_scaled,
                   xend = CTI_sens, yend = CPI_sens, color = "b4_cfc_col"),
               arrow = arrow(length = unit(0.1, "inches"))) +
  geom_point(data=b4_cfc_max,
             aes(x = CTI_ambient_scaled, y = CPI_ambient_scaled), color = "black") +
  geom_point(data=b4_cfc_max,
             aes(x = CTI_sens, y = CPI_sens), color = "red") +
  geom_segment(data=b4_hwrc_max,
               aes(x = CTI_ambient_scaled, y = CPI_ambient_scaled,
                   xend = CTI_sens, yend = CPI_sens, color = "b4_hwrc_col"),
               arrow = arrow(length = unit(0.1, "inches"))) +
  geom_point(data=b4_hwrc_max,
             aes(x = CTI_ambient_scaled, y = CPI_ambient_scaled), color = "black") +
  geom_point(data=b4_hwrc_max,
             aes(x = CTI_sens, y = CPI_sens), color = "red") +
  labs(x = "CTI", y = "CPI") +
  scale_colour_manual(name="Experiment",values=cols) +
  theme_minimal()
```

### Treatment effect for the 7th year of each experiment (PHACE was the shortest duration experiment at 7 years long)

```{r, warning=F,message=F,fig.width = 8, fig.height = 5}
# Filter so all experiments contain only their 7th year of data
# PHACE ran the shortest at 7 years in length; looking at the 7th year for all
tera_seventh_year <- CTI_CPI_tera_sens %>%
  filter(year == 2018)
jrgce_seventh_year <- CTI_CPI_jrgce_sens %>%
  filter(year == 2005)
phace_seventh_year <- CTI_CPI_phace_sens %>%
  filter(year == 2013)
ok_seventh_year <- CTI_CPI_ok_sens %>%
  filter(year == 2006)
b4_cfc_seventh_year <- CTI_CPI_b4_sens_cfc %>%
  filter(year == 2014)
b4_hwrc_seventh_year <- CTI_CPI_b4_sens_hwrc %>%
  filter(year == 2014)

cols <- c("jrgce_col"="#332288","phace_col"="#117733","tera_col"="#882255",
         "ok_col" =  "#88CCEE", "b4_cfc_col" = "#AA4499", "b4_hwrc_col" = "#DDCC77")

ggplot() +
  geom_segment(data=tera_seventh_year,
               aes(x = CTI_ambient_scaled, y = CPI_ambient_scaled,
                   xend = CTI_sens, yend = CPI_sens, color = "tera_col"),
               arrow = arrow(length = unit(0.1, "inches"))) +
  geom_point(data=tera_seventh_year,
             aes(x = CTI_ambient_scaled, y = CPI_ambient_scaled), color = "black") +
  geom_point(data=tera_seventh_year,
             aes(x = CTI_sens, y = CPI_sens), color = "red") +
  geom_segment(data=phace_seventh_year,
               aes(x = CTI_ambient_scaled, y = CPI_ambient_scaled,
                   xend = CTI_sens, yend = CPI_sens, color = "phace_col"),
               arrow = arrow(length = unit(0.1, "inches"))) +
  geom_point(data=phace_seventh_year,
             aes(x = CTI_ambient_scaled, y = CPI_ambient_scaled), color = "black") +
  geom_point(data=phace_seventh_year,
             aes(x = CTI_sens, y = CPI_sens), color = "red") +
  geom_segment(data=jrgce_seventh_year,
               aes(x = CTI_ambient_scaled, y = CPI_ambient_scaled,
                   xend = CTI_sens, yend = CPI_sens, color = "jrgce_col"),
               arrow = arrow(length = unit(0.1, "inches"))) +
  geom_point(data=jrgce_seventh_year,
             aes(x = CTI_ambient_scaled, y = CPI_ambient_scaled), color = "black") +
  geom_point(data=jrgce_seventh_year,
             aes(x = CTI_sens, y = CPI_sens), color = "red") +
   geom_segment(data=ok_seventh_year,
               aes(x = CTI_ambient_scaled, y = CPI_ambient_scaled,
                   xend = CTI_sens, yend = CPI_sens, color = "ok_col"),
               arrow = arrow(length = unit(0.1, "inches"))) +
  geom_point(data=ok_seventh_year,
             aes(x = CTI_ambient_scaled, y = CPI_ambient_scaled), color = "black") +
  geom_point(data=ok_seventh_year,
             aes(x = CTI_sens, y = CPI_sens), color = "red") +
  geom_segment(data=b4_cfc_seventh_year,
               aes(x = CTI_ambient_scaled, y = CPI_ambient_scaled,
                   xend = CTI_sens, yend = CPI_sens, color = "b4_cfc_col"),
               arrow = arrow(length = unit(0.1, "inches"))) +
  geom_point(data=b4_cfc_seventh_year,
             aes(x = CTI_ambient_scaled, y = CPI_ambient_scaled), color = "black") +
  geom_point(data=b4_cfc_seventh_year,
             aes(x = CTI_sens, y = CPI_sens), color = "red") +
  geom_segment(data=b4_hwrc_seventh_year,
               aes(x = CTI_ambient_scaled, y = CPI_ambient_scaled,
                   xend = CTI_sens, yend = CPI_sens, color = "b4_hwrc_col"),
               arrow = arrow(length = unit(0.1, "inches"))) +
  geom_point(data=b4_hwrc_seventh_year,
             aes(x = CTI_ambient_scaled, y = CPI_ambient_scaled), color = "black") +
  geom_point(data=b4_hwrc_seventh_year,
             aes(x = CTI_sens, y = CPI_sens), color = "red") +
  labs(x = "CTI", y = "CPI") +
  scale_colour_manual(name="Experiment",values=cols) +
  theme_minimal()
```

## Species-level analyses

### Boxplots of species' temperature niches that showed decreases, increases, or no change as a function of increasing CTI

```{r,message=F,warning=F,fig.width = 15, fig.height = 6}
# Calculate mean spp abundance per year and treatment
phace_test <- phace %>%
  group_by(year, temp_treatment,species) %>%
  summarize(avg_abun = mean(rel_abun, na.rm=T))
ok_test <- ok %>%
  group_by(year, temp_treatment,species) %>%
  summarize(avg_abun = mean(rel_abun, na.rm=T))
jrgce_test <- jrgce %>%
  group_by(year, temp_treatment,species) %>%
  summarize(avg_abun = mean(percent_cover, na.rm=T))
tera_test <- tera %>%
  filter(!(species == "Total Planted Species")) %>%
  group_by(year, temp_treatment,species) %>%
  summarize(avg_abun = mean(percent_cover, na.rm=T))
#b4_3.4 <- b4 %>%
#  filter(temp_treatment == "3.4" | temp_treatment == "amb")
b4_cfc <- b4 %>%
  filter(site == "CFC")
b4_hwrc <- b4 %>%
  filter(site == "HWRC")
b4_test_cfc <- b4_cfc %>%
  group_by(year, temp_treatment,species) %>%
  summarize(avg_abun = mean(rel_abun, na.rm=T))
b4_test_hwrc <- b4_hwrc %>%
  group_by(year, temp_treatment,species) %>%
  summarize(avg_abun = mean(rel_abun, na.rm=T))

# Calculating mean CTI per year and treatment
CTI_yearly_avg_phace <- CTI_phace %>%
  group_by(year,temp_treatment) %>%
  summarize(mean_CTI = mean(CTI))
CTI_yearly_avg_ok <- CTI_ok %>%
  group_by(year,temp_treatment) %>%
  summarize(mean_CTI = mean(CTI))
CTI_yearly_avg_jrgce <- CTI_jrgce %>%
  group_by(year,temp_treatment) %>%
  summarize(mean_CTI = mean(CTI))
CTI_yearly_avg_tera <- CTI_teracon %>%
  group_by(year,temp_treatment) %>%
  summarize(mean_CTI = mean(CTI))
#CTI_b4_3.4 <- CTI_b4 %>%
#  filter(temp_treatment == "3.4" | temp_treatment == "amb")
CTI_yearly_avg_b4_cfc <- CTI_b4_cfc %>%
  group_by(year,temp_treatment) %>%
  summarize(mean_CTI = mean(CTI))
CTI_yearly_avg_b4_hwrc <- CTI_b4_hwrc %>%
  group_by(year,temp_treatment) %>%
  summarize(mean_CTI = mean(CTI))

# Merging CTI and species abundance data
phace_spp_abun <- merge(CTI_yearly_avg_phace,phace_test, by=c("year","temp_treatment"))
ok_spp_abun <- merge(CTI_yearly_avg_ok,ok_test, by=c("year","temp_treatment"))
jrgce_spp_abun <- merge(CTI_yearly_avg_jrgce,jrgce_test, by=c("year","temp_treatment"))
tera_spp_abun <- merge(CTI_yearly_avg_tera,tera_test, by=c("year","temp_treatment"))
b4_spp_abun_cfc <- merge(CTI_yearly_avg_b4_cfc,b4_test_cfc, by=c("year","temp_treatment"))
b4_spp_abun_hwrc <- merge(CTI_yearly_avg_b4_hwrc,b4_test_hwrc, by=c("year","temp_treatment"))

### Stats: Which species show significant increases or decreases in abundance as a function of CTI?
# Perform regression analysis for each species
results_phace <- phace_spp_abun %>%
  group_by(species) %>%
  do(model = lm(avg_abun ~ mean_CTI, data = .))
results_ok <- ok_spp_abun %>%
  group_by(species) %>%
  do(model = lm(avg_abun ~ mean_CTI, data = .))
results_jrgce <- jrgce_spp_abun %>%
  filter(!(species == "Madia gracilis" | species == "Quercus lobata" | species == "Festuca DUMMY")) %>% # removing spp w/ NAs
  group_by(species) %>%
  do(model = lm(avg_abun ~ mean_CTI, data = .))
results_tera <- tera_spp_abun %>%
  filter(!(species == "Total Planted Species")) %>%
  group_by(species) %>%
  do(model = lm(avg_abun ~ mean_CTI, data = .))
results_b4_cfc <- b4_spp_abun_cfc %>%
  group_by(species) %>%
  do(model = lm(avg_abun ~ mean_CTI, data = .))
results_b4_hwrc <- b4_spp_abun_hwrc %>%
  group_by(species) %>%
  do(model = lm(avg_abun ~ mean_CTI, data = .))

# Extract coefficients and p-values
results_phace_summary <- results_phace %>%
  rowwise() %>%
  mutate(slope = coef(model)[["mean_CTI"]],
         p_value = summary(model)$coefficients["mean_CTI", "Pr(>|t|)"])
results_ok_summary <- results_ok %>%
  rowwise() %>%
  mutate(slope = coef(model)[["mean_CTI"]],
         p_value = summary(model)$coefficients["mean_CTI", "Pr(>|t|)"])
results_jrgce_summary <- results_jrgce %>%
  rowwise() %>%
  mutate(slope = coef(model)[["mean_CTI"]],
         p_value = summary(model)$coefficients["mean_CTI", "Pr(>|t|)"])
results_tera_summary <- results_tera %>%
  rowwise() %>%
  mutate(slope = coef(model)[["mean_CTI"]],
         p_value = summary(model)$coefficients["mean_CTI", "Pr(>|t|)"])
results_b4_summary_cfc <- results_b4_cfc %>%
  rowwise() %>%
  mutate(slope = coef(model)[["mean_CTI"]],
         p_value = summary(model)$coefficients["mean_CTI", "Pr(>|t|)"])
results_b4_summary_hwrc <- results_b4_hwrc %>%
  rowwise() %>%
  mutate(slope = coef(model)[["mean_CTI"]],
         p_value = summary(model)$coefficients["mean_CTI", "Pr(>|t|)"])

# Apply multiple testing correction (Benjamini-Hochberg)
results_phace_summary <- results_phace_summary %>%
  mutate(adj_p_value = p.adjust(p_value, method = "BH"))
results_ok_summary <- results_ok_summary %>%
  mutate(adj_p_value = p.adjust(p_value, method = "BH"))
results_jrgce_summary <- results_jrgce_summary %>%
  mutate(adj_p_value = p.adjust(p_value, method = "BH"))
results_tera_summary <- results_tera_summary %>%
  mutate(adj_p_value = p.adjust(p_value, method = "BH"))
results_b4_summary_cfc <- results_b4_summary_cfc %>%
  mutate(adj_p_value = p.adjust(p_value, method = "BH"))
results_b4_summary_hwrc <- results_b4_summary_hwrc %>%
  mutate(adj_p_value = p.adjust(p_value, method = "BH"))

# Filter out species with significant effects
results_phace_sig <- results_phace_summary %>%
  filter(adj_p_value <= 0.06)
results_ok_sig <- results_ok_summary %>%
  filter(adj_p_value <= 0.06)
results_jrgce_sig <- results_jrgce_summary %>%
  filter(adj_p_value <= 0.06)
results_tera_sig <- results_tera_summary %>%
  filter(adj_p_value <= 0.06)
results_b4_sig_cfc <- results_b4_summary_cfc %>%
  filter(adj_p_value <= 0.06)
results_b4_sig_hwrc <- results_b4_summary_hwrc %>%
  filter(adj_p_value <= 0.06)

# Filtering data to sig. species
phace_spp_list <- unique(results_phace_sig$species)
phace_sig_spp <- phace_spp_abun %>%
  filter(species %in% phace_spp_list)

ok_spp_list <- unique(results_ok_sig$species)
ok_sig_spp <- ok_spp_abun %>%
  filter(species %in% ok_spp_list)

jrgce_spp_list <- unique(results_jrgce_sig$species)
jrgce_sig_spp <- jrgce_spp_abun %>%
  filter(species %in% jrgce_spp_list)

tera_spp_list <- unique(results_tera_sig$species)
tera_sig_spp <- tera_spp_abun %>%
  filter(species %in% tera_spp_list)

b4_spp_list_cfc <- unique(results_b4_sig_cfc$species)
b4_sig_spp_cfc <- b4_spp_abun_cfc %>%
  filter(species %in% b4_spp_list_cfc)

b4_spp_list_hwrc <- unique(results_b4_sig_hwrc$species)
b4_sig_spp_hwrc <- b4_spp_abun_hwrc %>%
  filter(species %in% b4_spp_list_hwrc)

### Fig: boxplot of spp that increase, decrease, or show no change 
phace_merge <- merge(phace_test,niche_est_phace_avg, by="species")
phace_merge_stats <- merge(phace_merge,results_phace_summary, by="species")
phace_merge_stats$significance <- ifelse(phace_merge_stats$adj_p_value <= 0.06, "significant", "not significant")
phace_merge_stats$slope_value <- ifelse(phace_merge_stats$slope > 0, "positive", "negative")
phace_merge_stats$change <- ifelse(phace_merge_stats$significance == "significant" & phace_merge_stats$slope_value == "positive", "increase",
                                    ifelse(phace_merge_stats$significance == "significant" & phace_merge_stats$slope_value == "negative", "decrease", "no change"))
phace_merge_stats <- phace_merge_stats %>%
  filter(!is.na(change))

ok_merge <- merge(ok_test,niche_est_ok_avg, by="species")
ok_merge_stats <- merge(ok_merge,results_ok_summary, by="species")
ok_merge_stats$significance <- ifelse(ok_merge_stats$adj_p_value <= 0.06, "significant", "not significant")
ok_merge_stats$slope_value <- ifelse(ok_merge_stats$slope > 0, "positive", "negative")
ok_merge_stats$change <- ifelse(ok_merge_stats$significance == "significant" & ok_merge_stats$slope_value == "positive", "increase",
                                   ifelse(ok_merge_stats$significance == "significant" & ok_merge_stats$slope_value == "negative", "decrease", "no change"))
ok_merge_stats <- ok_merge_stats %>%
  filter(!is.na(change))

tera_merge <- merge(tera_test,niche_est_tera_avg, by="species")
tera_merge_stats <- merge(tera_merge,results_tera_summary, by="species")
tera_merge_stats$significance <- ifelse(tera_merge_stats$adj_p_value <= 0.06, "significant", "not significant")
tera_merge_stats$slope_value <- ifelse(tera_merge_stats$slope > 0, "positive", "negative")
tera_merge_stats$change <- ifelse(tera_merge_stats$significance == "significant" & tera_merge_stats$slope_value == "positive", "increase",
                                   ifelse(tera_merge_stats$significance == "significant" & tera_merge_stats$slope_value == "negative", "decrease", "no change"))

jrgce_merge <- merge(jrgce_test,niche_est_jrgce_avg, by="species")
jrgce_merge_stats <- merge(jrgce_merge,results_jrgce_summary, by="species")
jrgce_merge_stats$significance <- ifelse(jrgce_merge_stats$adj_p_value <= 0.06, "significant", "not significant")
jrgce_merge_stats$slope_value <- ifelse(jrgce_merge_stats$slope > 0, "positive", "negative")
jrgce_merge_stats$change <- ifelse(jrgce_merge_stats$significance == "significant" & jrgce_merge_stats$slope_value == "positive", "increase",
                                  ifelse(jrgce_merge_stats$significance == "significant" & jrgce_merge_stats$slope_value == "negative", "decrease", "no change"))
jrgce_merge_stats <- jrgce_merge_stats %>%
  filter(!is.na(change))

b4_merge_cfc <- merge(b4_test_cfc,niche_est_b4_avg_cfc, by="species")
b4_merge_stats_cfc <- merge(b4_merge_cfc,results_b4_summary_cfc, by="species")
b4_merge_stats_cfc$significance <- ifelse(b4_merge_stats_cfc$adj_p_value <= 0.06, "significant", "not significant")
b4_merge_stats_cfc$slope_value <- ifelse(b4_merge_stats_cfc$slope > 0, "positive", "negative")
b4_merge_stats_cfc$change <- ifelse(b4_merge_stats_cfc$significance == "significant" & b4_merge_stats_cfc$slope_value == "positive", "increase",
                                   ifelse(b4_merge_stats_cfc$significance == "significant" & b4_merge_stats_cfc$slope_value == "negative", "decrease", "no change"))
b4_merge_stats_cfc <-  b4_merge_stats_cfc %>%
  filter(!is.na(change))

b4_merge_hwrc <- merge(b4_test_hwrc,niche_est_b4_avg_hwrc, by="species")
b4_merge_stats_hwrc <- merge(b4_merge_hwrc,results_b4_summary_hwrc, by="species")
b4_merge_stats_hwrc$significance <- ifelse(b4_merge_stats_hwrc$adj_p_value <= 0.06, "significant", "not significant")
b4_merge_stats_hwrc$slope_value <- ifelse(b4_merge_stats_hwrc$slope > 0, "positive", "negative")
b4_merge_stats_hwrc$change <- ifelse(b4_merge_stats_hwrc$significance == "significant" & b4_merge_stats_hwrc$slope_value == "positive", "increase",
                                    ifelse(b4_merge_stats_hwrc$significance == "significant" & b4_merge_stats_hwrc$slope_value == "negative", "decrease", "no change"))
b4_merge_stats_hwrc <-  b4_merge_stats_hwrc %>%
  filter(!is.na(change))

# Code to determine median niche for each community to color species by
#median(niche_est_tera_avg$temp_niche) # 8.75
#median(niche_est_phace_avg$temp_niche) # 8.65
#median(niche_est_ok_avg$temp_niche) # 14.78
#median(niche_est_jrgce_avg$temp_niche) # 14.65
#median(niche_est_b4_avg_cfc$temp_niche) # 7.95
#median(niche_est_b4_avg_hwrc$temp_niche) # 7.84

tera_merge_stats$niche_cat <- ifelse(tera_merge_stats$temp_niche >= 8.75, "High temp niche", "Low temp niche")
phace_merge_stats$niche_cat <- ifelse(phace_merge_stats$temp_niche >= 8.65, "High temp niche", "Low temp niche")
ok_merge_stats$niche_cat <- ifelse(ok_merge_stats$temp_niche >= 14.78, "High temp niche", "Low temp niche")
jrgce_merge_stats$niche_cat <- ifelse(jrgce_merge_stats$temp_niche >= 14.65, "High temp niche", "Low temp niche")
b4_merge_stats_cfc$niche_cat <- ifelse(b4_merge_stats_cfc$temp_niche >= 7.95, "High temp niche", "Low temp niche")
b4_merge_stats_hwrc$niche_cat <- ifelse(b4_merge_stats_hwrc$temp_niche >= 7.84, "High temp niche", "Low temp niche")

tera_sig_spp_cat <- left_join(tera_sig_spp,tera_merge_stats,by=c("species","year","temp_treatment"))
phace_sig_spp_cat <- left_join(phace_sig_spp,phace_merge_stats,by=c("species","year","temp_treatment"))
ok_sig_spp_cat <- left_join(ok_sig_spp,ok_merge_stats,by=c("species","year","temp_treatment"))
jrgce_sig_spp_cat <- left_join(jrgce_sig_spp,jrgce_merge_stats,by=c("species","year","temp_treatment"))
b4_sig_spp_cat_cfc <- left_join(b4_sig_spp_cfc,b4_merge_stats_cfc,by=c("species","year","temp_treatment"))
b4_sig_spp_cat_hwrc <- left_join(b4_sig_spp_hwrc,b4_merge_stats_hwrc,by=c("species","year","temp_treatment"))


phace_sig <- ggplot(phace_merge_stats, aes(x = change, y = temp_niche)) +
  geom_boxplot() +
  labs(y = "Temperature niche", x = "Species abundance change as a function of increasing CTI",title="PHACE") +
  #facet_wrap(.~temp_treatment) +
  theme_bw()
ok_sig <- ggplot(ok_merge_stats, aes(x = change, y = temp_niche)) +
  geom_boxplot() +
  labs(y = "Temperature niche", x = "Species abundance change as a function of increasing CTI",title="Oklahoma") +
  #facet_wrap(.~temp_treatment) +
  theme_bw()
tera_sig <- ggplot(tera_merge_stats, aes(x = change, y = temp_niche)) +
  geom_boxplot() +
  labs(y = "Temperature niche", x = "Species abundance change as a function of increasing CTI",title="TeRaCON") +
  #facet_wrap(.~temp_treatment) +
  theme_bw()
jrgce_sig <- ggplot(jrgce_merge_stats, aes(x = change, y = temp_niche)) +
  geom_boxplot() +
  labs(y = "Temperature niche", x = "Species abundance change as a function of increasing CTI",title="JRGCE") +
  #facet_wrap(.~temp_treatment) +
  theme_bw()
b4_sig_cfc <- ggplot(b4_merge_stats_cfc, aes(x = change, y = temp_niche)) +
  geom_boxplot() +
  labs(y = "Temperature niche", x = "Species abundance change as a function of increasing CTI",title="B4Warmed CFC") +
  #facet_wrap(.~temp_treatment) +
  theme_bw()
b4_sig_hwrc <- ggplot(b4_merge_stats_hwrc, aes(x = change, y = temp_niche)) +
  geom_boxplot() +
  labs(y = "Temperature niche", x = "Species abundance change as a function of increasing CTI",title="B4Warmed HWRC") +
  #facet_wrap(.~temp_treatment) +
  theme_bw()

ggarrange(jrgce_sig,phace_sig,tera_sig,ok_sig,b4_sig_cfc,b4_sig_hwrc,
          nrow=2,ncol=3)
```

### Species that demonstrate a significant relationship with CTI from each experiment. Species are ordered & colored in terms of increasing thermal niches

```{r,message=F,warning=F,fig.width = 13, fig.height = 6}
# Plotting changes in abundance for significant species  
ggplot(jrgce_sig_spp_cat, aes(x = mean_CTI, y = avg_abun.x, color = temp_niche)) +
  geom_smooth(method="lm") +
  geom_point() +
  theme_minimal() +
  labs(title = "JRGCE",
       x = "Mean CTI",
       y = "Abundance (%)") +
  facet_wrap(~reorder(species, temp_niche), scales = "free_y") +
  scale_color_viridis_c(option = "viridis")

ggplot(phace_sig_spp_cat, aes(x = mean_CTI, y = avg_abun.x*100, color = temp_niche)) +
  geom_smooth(method="lm") +
  geom_point() +
  theme_minimal() +
  labs(title = "PHACE",
       x = "Mean CTI",
       y = "Abundance (%)") +
  facet_wrap(~reorder(species, temp_niche), scales = "free_y") +
  scale_color_viridis_c(option = "viridis")

ggplot(ok_sig_spp_cat, aes(x = mean_CTI, y = avg_abun.x*100, color = temp_niche)) +
  geom_smooth(method="lm") +
  geom_point() +
  theme_minimal() +
  labs(title = "Oklahoma",
       x = "Mean CTI",
       y = "Abundance (%)") +
  facet_wrap(~reorder(species, temp_niche), scales = "free_y") +
  scale_color_viridis_c(option = "viridis")

ggplot(tera_sig_spp_cat, aes(x = mean_CTI, y = avg_abun.x, color = temp_niche)) +
  geom_smooth(method="lm") +
  geom_point() +
  theme_minimal() +
  labs(title = "TeRaCON",
       x = "Mean CTI",
       y = "Abundance (%)") +
  facet_wrap(~reorder(species, temp_niche), scales = "free_y") +
  scale_color_viridis_c(option = "viridis")

ggplot(b4_sig_spp_cat_cfc, aes(x = mean_CTI, y = avg_abun.x*100, color = temp_niche)) +
  geom_smooth(method="lm") +
  geom_point() +
  theme_minimal() +
  labs(title = "B4Warmed CFC",
       x = "Mean CTI",
       y = "Abundance (%)") +
  facet_wrap(~reorder(species, temp_niche), scales = "free_y") +
  scale_color_viridis_c(option = "viridis")

ggplot(b4_sig_spp_cat_hwrc, aes(x = mean_CTI, y = avg_abun.x*100, color = temp_niche)) +
  geom_smooth(method="lm") +
  geom_point() +
  theme_minimal() +
  labs(title = "B4Warmed HWRC",
       x = "Mean CTI",
       y = "Abundance (%)") +
  facet_wrap(~reorder(species, temp_niche), scales = "free_y") +
  scale_color_viridis_c(option = "viridis")
```

### All species from each experiment on the same figure; better shows the dominance of particular species in terms of relative abundance

```{r,message=F,warning=F,fig.width = 13, fig.height = 6}
ggplot(jrgce_sig_spp_cat, aes(x = mean_CTI, y = avg_abun.x, group = species, color = temp_niche)) +
  geom_smooth(method="lm") +
  geom_point() +
  theme_minimal() +
  labs(title = "JRGCE",
       x = "Mean CTI",
       y = "Abundance (%)") +
  #facet_wrap(~reorder(species, temp_niche), scales = "free_y") +
  scale_color_viridis_c(option = "viridis")

ggplot(phace_sig_spp_cat, aes(x = mean_CTI, y = avg_abun.x*100, group = species,color = temp_niche)) +
  geom_smooth(method="lm") +
  geom_point() +
  theme_minimal() +
  labs(title = "PHACE",
       x = "Mean CTI",
       y = "Abundance (%)") +
  #facet_wrap(~reorder(species, temp_niche), scales = "free_y") +
  scale_color_viridis_c(option = "viridis")

ggplot(ok_sig_spp_cat, aes(x = mean_CTI, y = avg_abun.x*100, group = species,color = temp_niche)) +
  geom_smooth(method="lm") +
  geom_point() +
  theme_minimal() +
  labs(title = "Oklahoma",
       x = "Mean CTI",
       y = "Abundance (%)") +
  #facet_wrap(~reorder(species, temp_niche), scales = "free_y") +
  scale_color_viridis_c(option = "viridis")

ggplot(tera_sig_spp_cat, aes(x = mean_CTI, y = avg_abun.x, group = species, color = temp_niche)) +
  geom_smooth(method="lm") +
  geom_point() +
  theme_minimal() +
  labs(title = "TeRaCON",
       x = "Mean CTI",
       y = "Abundance (%)") +
  #facet_wrap(~reorder(species, temp_niche), scales = "free_y") +
  scale_color_viridis_c(option = "viridis")

ggplot(b4_sig_spp_cat_cfc, aes(x = mean_CTI, y = avg_abun.x*100, group = species, color = temp_niche)) +
  geom_smooth(method="lm") +
  geom_point() +
  theme_minimal() +
  labs(title = "B4Warmed CFC",
       x = "Mean CTI",
       y = "Abundance (%)") +
  #facet_wrap(~reorder(species, temp_niche), scales = "free_y") +
  scale_color_viridis_c(option = "viridis")

ggplot(b4_sig_spp_cat_hwrc, aes(x = mean_CTI, y = avg_abun.x*100, group = species, color = temp_niche)) +
  geom_smooth(method="lm") +
  geom_point() +
  theme_minimal() +
  labs(title = "B4Warmed HWRC",
       x = "Mean CTI",
       y = "Abundance (%)") +
  #facet_wrap(~reorder(species, temp_niche), scales = "free_y") +
  scale_color_viridis_c(option = "viridis")
```

## Plant traits

### Plant trait values as a function of the species' thermal niche (Trait data from TRY)

```{r,warning=F,message=F,fig.width = 13, fig.height = 6}
##### Fig: traits vs. temp niche
# make a dataframe containing all species and their temp niches
tera_niche_filt <- niche_est_tera %>%
  select(species,temp_niche) %>%
  distinct()
jrgce_niche_filt <- niche_est_jrgce %>%
  select(species,temp_niche) %>%
  distinct()
phace_niche_filt <- niche_est_phace %>%
  select(species,temp_niche) %>%
  distinct()
b4_cfc_niche_filt <- niche_est_b4 %>%
  filter(site == "CFC") %>%
  select(species,temp_niche) %>%
  distinct()
b4_hwrc_niche_filt <- niche_est_b4 %>%
  filter(site == "HWRC") %>%
  select(species,temp_niche) %>%
  distinct()
ok_niche_filt <- niche_est_ok %>%
  select(species,temp_niche) %>%
  distinct()

# Combine all species and temp niches into one dataframe
all_niche <- rbind(tera_niche_filt,jrgce_niche_filt,phace_niche_filt,b4_cfc_niche_filt,b4_hwrc_niche_filt,ok_niche_filt)

# Renaming species to match the names found in Try
all_niche$species[all_niche$species == "Agropyron repens"] <- "Elymus repens"
all_niche$species[all_niche$species == "Lysimachia arvensis"] <- "Anagallis arvensis"
all_niche$species[all_niche$species == "Stipa pulchra"] <- "Nassella pulchra"
all_niche$species[all_niche$species == "Andropogon gerardi"] <- "Andropogon gerardii"
all_niche$species[all_niche$species == "Koeleria cristata"] <- "Koeleria macrantha"
all_niche$species[all_niche$species == "Festuca perennis"] <- "Lolium perenne"
all_niche$species[all_niche$species == "Taraxia ovata"] <- "Camissonia ovata"
all_niche$species[all_niche$species == "Festuca bromoides"] <- "Vulpia bromoides"
all_niche$species[all_niche$species == "Logfia gallica"] <- "Filago gallica"
all_niche$species[all_niche$species == "Pascopyrum smithii"] <- "Elymus smithii"
all_niche$species[all_niche$species == "Hesperostipa comata"] <- "Stipa comata"                       
all_niche$species[all_niche$species == "Picradeniopsis oppositifolia"] <- "Bahia oppositifolia"
all_niche$species[all_niche$species == "Heterotheca villosa"] <- "Chrysopsis villosa"
all_niche$species[all_niche$species == "Cryptantha thyrsiflora"] <- "Oreocarya thyrsiflora"
all_niche$species[all_niche$species == "Cryptantha cinerea"] <- "Oreocarya suffruticosa"
all_niche$species[all_niche$species == "Draba reptans"] <- "Tomostima reptans"
all_niche$species[all_niche$species == "Machaeranthera pinnatifida"] <- "Xanthisma spinulosum"
all_niche$species[all_niche$species == "Lesquerella montana"] <- "Physaria montana"
all_niche$species[all_niche$species == "Machaeranthera canescens"] <- "Dieteria canescens"
all_niche$species[all_niche$species == "Salsola tragus"] <- "Kali tragus"
all_niche$species[all_niche$species == "Populus tremouloides"] <- "Populus tremuloides"
all_niche$species[all_niche$species == "Calylophus serrulatus"] <- "Oenothera serrulata"
all_niche$species[all_niche$species == "Coreopsis grandiflora"] <- "Bidens sweetiana"
all_niche$species[all_niche$species == "Dichanthelium oligosanthes"] <- "Panicum oligosanthes"
all_niche$species[all_niche$species == "Gaura parviflora"] <- "Oenothera curtiflora"
all_niche$species[all_niche$species == "Hedyotis nigricans"] <- "Stenaria nigricans"
all_niche$species[all_niche$species == "Psoralidium tenuiflorum"] <- "Pediomelum tenuiflorum"
all_niche$species[all_niche$species == "Solidago ludiviciana"] <- "Solidago ludoviciana"
all_niche$species[all_niche$species == "Stenosiphon linifolius"] <- "Oenothera glaucifolia"
all_niche$species[all_niche$species == "Symphyotrichum ericoides "] <- "Symphyotrichum ericoides"
all_niche$species[all_niche$species == "Amphiachyris dracunculoides"] <- "Gutierrezia dracunculoides"
all_niche$species[all_niche$species == "Diodia teres"] <- "Hexasepalum teres"
all_niche$species[all_niche$species == "Stipa tenuissima"] <- "Nassella tenuissima"
all_niche$species[all_niche$species == "Rubus calycinoides"] <- "Rubus rolfei"
all_niche$species[all_niche$species == "Rubus pentalobus"] <- "Rubus rolfei"
all_niche$species[all_niche$species == "Ruellia hirta"] <- "Strobilanthes glutinosus"
all_niche$species[all_niche$species == "solidago drummondii"] <- "Solidago drummondii"
all_niche$species[all_niche$species == "Symphyotrichum ontarionsis"] <- "Symphyotrichum ontarionis"
all_niche$species[all_niche$species == "Haplopappus ciliatus"] <- "Grindelia ciliata"
all_niche$species[all_niche$species == "Heterotheca latifolia"] <- "Heterotheca subaxillaris"
all_niche$species[all_niche$species == "Muhlenbergia capillaris "] <- "Muhlenbergia capillaris"
all_niche$species[all_niche$species == "Setaria glauca"] <- "Pennisetum glaucum"
all_niche$species[all_niche$species == "Ambrosia trida"] <- "Ambrosia trifida"
all_niche$species[all_niche$species == "Diodia ocymifolia"] <- "Spermacoce ocymifolia"
all_niche$species[all_niche$species == "Grindelia papposa"] <- "Grindelia ciliata"
all_niche$species[all_niche$species == "Tragia cannabina"] <- "Tragia plukenetii"
all_niche$species[all_niche$species == "Trifolium purpureum "] <- "Trifolium purpureum"
all_niche$species[all_niche$species == "Melilotus alba"] <- "Trigonella alba"
all_niche$species[all_niche$species == "Coreopsis tinctoria"] <- "Bidens tinctoria"

# Combine all species niche data with trait data
# rename AccSpeciesName column in traits dataframe
colnames(traits)[1] <- "species"
traits_niche <- left_join(all_niche,traits, by="species")
traits_niche <- traits_niche %>%
  filter(!is.na(TraitID))

# Fixing long trait names
traits_niche$TraitName[traits_niche$TraitName == "Leaf dry mass per leaf fresh mass (leaf dry matter content, LDMC)"] <- "LDMC"
traits_niche$TraitName[traits_niche$TraitName == "Fine root length per fine root dry mass (specific fine root length, SRL)"] <- "Specific fine root length"
traits_niche$TraitName[traits_niche$TraitName == "Root length per root dry mass (specific root length, SRL)"] <- "Specific root length"
traits_niche$TraitName[traits_niche$TraitName == "Plant biomass and allometry: Fine root length per plant"] <- "Fine root length"
traits_niche$TraitName[traits_niche$TraitName == "Leaf area per leaf dry mass (specific leaf area, SLA or 1/LMA): undefined if petiole is in- or excluded"] <- "SLA"
traits_niche$TraitName[traits_niche$TraitName == "Coarse root length per coarse root dry mass (specific coarse root length, SRL)"] <- "Specific coarse root length"
traits_niche$TraitName[traits_niche$TraitName == "Plant biomass and allometry: Coarse root length per plant"] <- "Coarse root length"

# Removing some high outliers and traits with little data
traits_niche_rem <- traits_niche %>%
  filter(TraitName == "Plant height vegetative") %>%
  filter(mean_trait_val < 400)
traits_niche_rem2 <- traits_niche %>%
  filter(TraitName == "Specific fine root length") %>%
  filter(mean_trait_val < 8000)
rest_of_data <- traits_niche %>%
  filter(!(TraitName == "Plant height vegetative" |
             TraitName == "Specific fine root length"))
combined_data <- bind_rows(traits_niche_rem, traits_niche_rem2, rest_of_data)
combined_data <- combined_data %>%
  filter(!(TraitName == "Fine root length" |
             TraitName == "Coarse root length" |
             TraitName == "Specific coarse root length" |
             TraitName == "Dispersal syndrome"))

# Make a function to select a given trait name and plot it against temp niche
trait_niche_plot <- function(data,trait){
  plot_data <- data %>%
    filter(TraitID == trait)
  
  trait_name <- plot_data$TraitName[1]
  
  ggplot(plot_data, aes(x = temp_niche, y = mean_trait_val)) +
    geom_point() +
    geom_smooth() +
    labs(title = trait_name, x = "Temperature niche (°C)", y = "Mean trait value per species") +
    theme_minimal()
}
#trait_niche_plot(traits_niche_rem,3106)

# Initialize a list to store plots
plots_list <- list()

# Check unique TraitID values
unique_trait_ids <- unique(combined_data$TraitID)

# Generate plots and store them in the list
for (trait in unique_trait_ids) {
  # Ensure `trait` is always a single entity
  if (length(trait) == 1) {
    # Assuming trait_niche_plot returns a plot
    p <- trait_niche_plot(combined_data, trait)
    plots_list[[as.character(trait)]] <- p  # Convert to character to avoid unintended issues with numeric indices
  } else {
    message("Warning: Encountered unexpected multi-length trait identifier.")
  }
}

# Arrange all plots in a grid
grid.arrange(grobs = plots_list)
```

### An interactive plot that shows the species name for the niche x trait value plot

```{r,warning=F,message=F}
# Interactive plot
combined_data_plotly <- combined_data %>%
  filter(TraitName == "Plant height vegetative")
plot_ly(data=combined_data_plotly, x = ~temp_niche, y = ~mean_trait_val,
        type="scatter",mode="markers",color=~species)

# Generate a list of all unique trait names
trait_names <- unique(combined_data$TraitName)

# Create a master plot without specific initial data
p <- plot_ly(type = 'scatter', mode = 'markers')

# Add a trace for each (trait, species) combination
for (i in seq_along(trait_names)) {
  trait_data <- combined_data[combined_data$TraitName == trait_names[i], ]
  species_names <- unique(trait_data$species)
  
  for (species in species_names) {
    species_data <- trait_data[trait_data$species == species, ]
    
    p <- add_trace(p,
                   data = species_data,
                   x = ~temp_niche,
                   y = ~mean_trait_val,
                   name = species,
                   text = ~paste("Species:", species),  # Custom hover text
                   hoverinfo = "text",  # Ensure that only custom text is displayed
                   colors = "Dark2",
                   visible = i == 1  # Show only the first trait initially
    )
  }
}

# Add the dropdown to switch visibility between traces
p <- layout(p,
            updatemenus = list(
              list(
                type = 'dropdown',
                active = 0,  # Initial active trait (0-indexed)
                buttons = lapply(seq_along(trait_names), function(i) {
                  list(
                    label = trait_names[i],
                    method = "update",
                    args = list(
                      list(visible = as.logical(rep(seq_along(trait_names) == i, each = length(species_names)))),
                      list(title = paste("Mean Trait Value for", trait_names[i]))  # Update the plot title based on the selected trait
                    )
                  )
                })
              )
            ),
            xaxis = list(title = "Temperature Niche"),
            yaxis = list(title = "Mean Trait Value")
)

p
```
