---
title: "EcoAcc - Experiments"
author: "Kara Dobson"
date: 2024-11-21
format: html
editor: visual
---

## Ecological Acclimation: Warming Experiments

```{r, echo = F, message = F, warning = F}
# Load packages
library(tidyverse)
library(ggpubr)

# Set path to turbo to get data
path_data = "/nfs/turbo/seas-zhukai/proj-ecoacc/TeRaCON/"
setwd(path_data)
# Load in data
# Data using global spp occurrences from GBIF + year-round MAT:
CTI_sens_teracon <- read.csv(" CTI_sens_teracon.csv")
CTI_teracon <- read.csv(" CTI_teracon.csv")
tera <- read.csv(" teracon_clean.csv")
NPP_teracon <- read.csv(" eco_response_teracon.csv")
NPP_overall_teracon <- read.csv(" eco_response_overall_teracon.csv")
CTI_CPI_teracon <- read.csv(" CTI_CPI_teracon.csv")
# Data using ecoregion 8 spp occurrences from GBIF + year-round MAT:
CTI_sens_teracon_lim <- read.csv(" CTI_sens_teracon_limited.csv")
CTI_teracon_lim <- read.csv(" CTI_teracon_limited.csv")
CTI_CPI_teracon_lim <- read.csv(" CTI_CPI_teracon_limited.csv")
# Data using ecoregion 8 spp occurrences from GBIF + 6-month temps (March-Aug):
CTI_sens_teracon_6month <- read.csv(" CTI_sens_6month_teracon_limited.csv")
CTI_teracon_6month <- read.csv(" CTI_6month_teracon_limited.csv")
CTI_CPI_teracon_6month <- read.csv(" CTI_CPI_6month_teracon_limited.csv")
# Data w/o big bluestem:
CTI_sens_teracon_noblue <- read.csv(" CTI_sens_teracon_nobluestem.csv")
CTI_teracon_noblue <- read.csv(" CTI_teracon_nobluestem.csv")
CTI_CPI_teracon_noblue <- read.csv(" CTI_CPI_teracon_nobluestem.csv")

# Set path to data
path_data = "/nfs/turbo/seas-zhukai/proj-ecoacc/JRGCE/"
setwd(path_data)
# Load in data
CTI_sens_jrgce <- read.csv(" CTI_sens_jrgce.csv")
CTI_jrgce <- read.csv(" CTI_jrgce.csv")
jrgce <- read.csv(" jrgce_clean.csv")
jrgce <- jrgce %>%
  mutate(temp_treatment = if_else(str_detect(treatment, "T"), "warmed", "ambient"))
NPP_jrgce <- read.csv(" eco_response_jrgce.csv")
NPP_overall_jrgce <- read.csv(" eco_response_overall_jrgce.csv")
CTI_CPI_jrgce <- read.csv(" CTI_CPI_jrgce.csv")

# Set path to data
path_data = "/nfs/turbo/seas-zhukai/proj-ecoacc/PHACE/"
setwd(path_data)
# Load in data
CTI_sens_phace <- read.csv(" CTI_sens_phace_limited.csv")
CTI_phace <- read.csv(" CTI_phace_limited.csv")
phace <- read.csv(" phace_clean.csv")
NPP_phace <- read.csv(" eco_response_phace.csv")
NPP_overall_phace <- read.csv(" eco_response_overall_phace.csv")
CTI_CPI_phace <- read.csv(" CTI_CPI_phace_limited.csv")
```

Using community composition and ecosystem response data from various warming experiments, we can visualize changes in community structure (e.g., CTI) alongside changes in function (e.g., ANPP) due to warming. Including ambient temperatures changes over the course of each experiment may also inform the observed responses.

```{r, message = F, warning = F, echo = F}
# CTI figures
CTI_tera_plot <- ggplot(CTI_sens_teracon_lim, aes(x = year, y = sensitivity)) +
  geom_smooth() +
  labs(x = "Year", y = "TeRaCON\nCTI (Warmed - Ambient)") +
  scale_x_continuous(breaks = seq(2012, 2023, by = 2)) +
  theme_bw()
CTI_jrgce_plot <- ggplot(CTI_sens_jrgce, aes(x = year, y = sensitivity)) +
  geom_smooth() +
  labs(x = "Year", y = "JRGCE\nCTI (Warmed - Ambient)") +
  scale_x_continuous(breaks = seq(1998, 2014, by = 3)) +
  theme_bw()
CTI_phace_plot <- ggplot(CTI_sens_phace, aes(x = year, y = sensitivity)) +
  geom_smooth() +
  labs(x = "Year", y = "PHACE\nCTI (Warmed - Ambient)") +
  scale_x_continuous(breaks = seq(2007, 2013)) +
  theme_bw()

# Biomass figures
NPP_teracon <- NPP_teracon %>%
  filter(variable == "mean_ab_bio")
npp_tera_plot <- ggplot(NPP_teracon, aes(x = year, y = sensitivity)) +
  geom_smooth() +
  labs(x = "Year", y = "Biomass (Warmed - Ambient)") +
  scale_x_continuous(breaks = seq(2012, 2023, by = 2)) +
  theme_bw()
npp_jrgce_plot <- ggplot(NPP_jrgce, aes(x = year, y = sensitivity)) +
  geom_smooth() +
  labs(x = "Year", y = "Biomass (Warmed - Ambient)") +
  scale_x_continuous(breaks = seq(1998, 2014, by = 3)) +
  theme_bw()
npp_phace_plot <- ggplot(NPP_phace, aes(x = year, y = sensitivity)) +
  geom_smooth() +
  labs(x = "Year", y = "Biomass (Warmed - Ambient)") +
  scale_x_continuous(breaks = seq(2007, 2013)) +
  theme_bw()

# Ambient temp change figures
temp_tera_plot <- ggplot(CTI_sens_teracon_lim, aes(x = year, y = mean_C_temp_summer)) +
  geom_smooth() +
  labs(x = "Year", y = "Ambient temperature (°C)") +
  scale_x_continuous(breaks = seq(2012, 2023, by = 2)) +
  theme_bw()
temp_jrgce_plot <- ggplot(CTI_sens_jrgce, aes(x = year, y = mean_C_temp_summer)) +
  geom_smooth() +
  labs(x = "Year", y = "Ambient temperature (°C)") +
  scale_x_continuous(breaks = seq(1998, 2014, by = 3)) +
  theme_bw()
```

### Changes in CTI, NPP, and ambient temperatures

```{r, echo = F, message = F, warning = F, fig.width = 10, fig.height = 7}
ggarrange(CTI_tera_plot,npp_tera_plot,temp_tera_plot,
          CTI_jrgce_plot,npp_jrgce_plot,temp_jrgce_plot,
          CTI_phace_plot,npp_phace_plot,
          ncol = 3, nrow=3, common.legend = T, legend = "right")
```

```{r, echo = F, message = F, warning = F}
##### Fig: correlating biomass w/ CTI
# TeRaCON
CTI_yearly_avg_tera <- CTI_teracon_lim %>%
  group_by(year,temp_treatment) %>%
  summarize(mean_CTI = mean(CTI))
NPP_CTI_teracon <- left_join(CTI_yearly_avg_tera, NPP_overall_teracon, by=c("year","temp_treatment"))
tera_scatter <- ggscatter(NPP_CTI_teracon, x = "mean_ab_bio", y = "mean_CTI", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Biomass", ylab = "TeRaCON\nCTI")
# JRGCE
CTI_yearly_avg_jrgce <- CTI_jrgce %>%
  group_by(year,temp_treatment) %>%
  summarize(mean_CTI = mean(CTI))
NPP_CTI_jrgce <- left_join(CTI_yearly_avg_jrgce, NPP_overall_jrgce, by=c("year","temp_treatment"))
jrgce_scatter <- ggscatter(NPP_CTI_jrgce, x = "mean_ab_bio", y = "mean_CTI", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Biomass", ylab = "JRGCE\nCTI")
# PHACE
CTI_yearly_avg_phace <- CTI_phace %>%
  group_by(year,temp_treatment) %>%
  summarize(mean_CTI = mean(CTI))
NPP_CTI_phace <- left_join(CTI_yearly_avg_phace, NPP_overall_phace, by=c("year","temp_treatment"))
phace_scatter <- ggscatter(NPP_CTI_phace, x = "mean_ab_bio", y = "mean_CTI", 
                           add = "reg.line", conf.int = TRUE, 
                           cor.coef = TRUE, cor.method = "pearson",
                           xlab = "Biomass", ylab = "PHACE\nCTI")
```

### Correlation between biomass and CTI

```{r, echo = F, message = F, warning = F, fig.width = 13, fig.height = 6}
ggarrange(tera_scatter,jrgce_scatter,phace_scatter,
          ncol=3)
```

```{r, echo=F,warning=F,message=F}
##### Fig: Arrows pointing from ambient to warmed for each year #####
arrow_teracon <- ggplot(CTI_CPI_teracon_lim) +
  geom_segment(aes(x = CTI_HTamb, y = CPI_HTamb, 
                   xend = CTI_HTelv, yend = CPI_HTelv,
                   color = year),
               arrow = arrow(length = unit(0.1, "inches"))) +
  geom_point(aes(x = CTI_HTamb, y = CPI_HTamb), color = "black") +
  geom_point(aes(x = CTI_HTelv, y = CPI_HTelv), color = "red") +
  labs(x = "CTI", y = "CPI", title = "TeRaCON") +
  scale_color_viridis_c(option = "magma") +
  theme_minimal()

arrow_jrgce <- ggplot(CTI_CPI_jrgce) +
  geom_segment(aes(x = CTI_ambient, y = CPI_ambient, 
                   xend = CTI_warmed, yend = CPI_warmed,
                   color = year),
               arrow = arrow(length = unit(0.1, "inches"))) +
  geom_point(aes(x = CTI_ambient, y = CPI_ambient), color = "black") +
  geom_point(aes(x = CTI_warmed, y = CPI_warmed), color = "red") +
  labs(x = "CTI", y = "CPI", title = "JRGCE") +
  scale_color_viridis_c(option = "magma") +
  theme_minimal()

arrow_phace <- ggplot(CTI_CPI_phace) +
  geom_segment(aes(x = CTI_ambient, y = CPI_ambient, 
                   xend = CTI_warmed, yend = CPI_warmed,
                   color = year),
               arrow = arrow(length = unit(0.1, "inches"))) +
  geom_point(aes(x = CTI_ambient, y = CPI_ambient), color = "black") +
  geom_point(aes(x = CTI_warmed, y = CPI_warmed), color = "red") +
  labs(x = "CTI", y = "CPI", title = "PHACE") +
  scale_color_viridis_c(option = "magma") +
  theme_minimal()
```

### Yearly shifts in CTI and CPI due to the warming treatment

```{r, echo=F,warning=F,message=F, fig.width = 13, fig.height = 6}
ggarrange(arrow_teracon,arrow_jrgce,arrow_phace,
          ncol = 3, nrow=1)
```

```{r, echo=F,message=F,warning=F}
##### Fig: Mean CTI over time in warmed and ambient #####
# note: could also change y = CTI to a different metric (CTI_sd, CTI_skew, etc.)
CTI_teracon_plot <- ggplot(CTI_teracon_lim, aes(x = year, y = CTI, color = temp_treatment, group=temp_treatment)) +
  geom_jitter(alpha = 0.1,
              position = position_jitterdodge(dodge.width = 0.7)) +  # Add jittered points
  stat_summary(fun = mean,
               fun.min = mean,
               fun.max = mean,
               geom = "line",
              # width = 0.4,
               position = position_dodge(width = 0.7),
               aes(color = temp_treatment, group = temp_treatment)) +
  labs(title="TeRaCON") +
  theme_minimal() +
  scale_color_manual(values = c("HTamb" = "blue", "HTelv" = "red"))

CTI_jrgce <- CTI_jrgce %>%
  filter(!(year == 1998))
CTI_jrgce_plot <- ggplot(CTI_jrgce, aes(x = year, y = CTI, color = temp_treatment, group=temp_treatment)) +
  geom_jitter(alpha = 0.1,
              position = position_jitterdodge(dodge.width = 0.7)) +  # Add jittered points
  stat_summary(fun = mean,
               fun.min = mean,
               fun.max = mean,
               geom = "line",
               #width = 0.4,
               position = position_dodge(width = 0.7),
               aes(color = temp_treatment, group = temp_treatment)) +
  labs(title = "JRGCE") +
  theme_minimal() +
  scale_color_manual(values = c("ambient" = "blue", "warmed" = "red"))

CTI_phace_plot <- ggplot(CTI_phace, aes(x = year, y = CTI, color = temp_treatment, group=temp_treatment)) +
  geom_jitter(alpha = 0.1,
              position = position_jitterdodge(dodge.width = 0.7)) +  # Add jittered points
  stat_summary(fun = mean,
               fun.min = mean,
               fun.max = mean,
               geom = "line",
               #width = 0.4,
               position = position_dodge(width = 0.7),
               aes(color = temp_treatment, group = temp_treatment)) +
  labs(title = "PHACE") +
  theme_minimal() +
  scale_color_manual(values = c("ambient" = "blue", "warmed" = "red"))
```

### Temporal trends in CTI between treatments

```{r, echo=F,warning=F,message=F,fig.width = 13, fig.height = 6}
ggarrange(CTI_teracon_plot,CTI_jrgce_plot,CTI_phace_plot,
          ncol = 3, nrow=1)
```

```{r,echo=F,warning=F,message=F}
# Calculate mean spp abundance per year and treatment
phace_test <- phace %>%
  group_by(year, temp_treatment,species) %>%
  summarize(avg_abun = mean(rel_abun))
jrgce_test <- jrgce %>%
  group_by(year, temp_treatment,species) %>%
  summarize(avg_abun = mean(percent_cover))
tera_test <- tera %>%
  group_by(year, temp_treatment,species) %>%
  summarize(avg_abun = mean(percent_cover, na.rm=T))

# Calculating mean CTI per year and treatment
CTI_yearly_avg_phace <- CTI_phace %>%
  group_by(year,temp_treatment) %>%
  summarize(mean_CTI = mean(CTI))
CTI_yearly_avg_jrgce <- CTI_jrgce %>%
  group_by(year,temp_treatment) %>%
  summarize(mean_CTI = mean(CTI))
CTI_yearly_avg_tera <- CTI_teracon_lim %>%
  group_by(year,temp_treatment) %>%
  summarize(mean_CTI = mean(CTI))

# Merging CTI and species abundance data
phace_spp_abun <- merge(CTI_yearly_avg_phace,phace_test, by=c("year","temp_treatment"))
jrgce_spp_abun <- merge(CTI_yearly_avg_jrgce,jrgce_test, by=c("year","temp_treatment"))
tera_spp_abun <- merge(CTI_yearly_avg_tera,tera_test, by=c("year","temp_treatment"))

### Stats: Which species show significant increases or decreases in abundance as a function of CTI?
# Perform regression analysis for each species
results_phace <- phace_spp_abun %>%
  group_by(species) %>%
  do(model = lm(avg_abun ~ mean_CTI, data = .))
results_jrgce <- jrgce_spp_abun %>%
  filter(!(species == "Madia gracilis" | species == "Quercus lobata" | species == "Festuca DUMMY")) %>% # removing spp w/ NAs
  group_by(species) %>%
  do(model = lm(avg_abun ~ mean_CTI, data = .))
results_tera <- tera_spp_abun %>%
  filter(!(species == "Total Planted Species")) %>%
  group_by(species) %>%
  do(model = lm(avg_abun ~ mean_CTI, data = .))

# Extract coefficients and p-values
results_phace_summary <- results_phace %>%
  rowwise() %>%
  mutate(slope = coef(model)[["mean_CTI"]],
         p_value = summary(model)$coefficients["mean_CTI", "Pr(>|t|)"])
results_jrgce_summary <- results_jrgce %>%
  rowwise() %>%
  mutate(slope = coef(model)[["mean_CTI"]],
         p_value = summary(model)$coefficients["mean_CTI", "Pr(>|t|)"])
results_tera_summary <- results_tera %>%
  rowwise() %>%
  mutate(slope = coef(model)[["mean_CTI"]],
         p_value = summary(model)$coefficients["mean_CTI", "Pr(>|t|)"])

# Apply multiple testing correction (Benjamini-Hochberg)
results_phace_summary <- results_phace_summary %>%
  mutate(adj_p_value = p.adjust(p_value, method = "BH"))
results_jrgce_summary <- results_jrgce_summary %>%
  mutate(adj_p_value = p.adjust(p_value, method = "BH"))
results_tera_summary <- results_tera_summary %>%
  mutate(adj_p_value = p.adjust(p_value, method = "BH"))

# Filter out species with significant effects
results_phace_sig <- results_phace_summary %>%
  filter(adj_p_value <= 0.06)
results_jrgce_sig <- results_jrgce_summary %>%
  filter(adj_p_value <= 0.06)
results_tera_sig <- results_tera_summary %>%
  filter(adj_p_value <= 0.06)

# Filtering data to sig. species
phace_spp_list <- unique(results_phace_sig$species)
phace_sig_spp <- phace_spp_abun %>%
  filter(species %in% phace_spp_list)

jrgce_spp_list <- unique(results_jrgce_sig$species)
jrgce_sig_spp <- jrgce_spp_abun %>%
  filter(species %in% jrgce_spp_list)

tera_spp_list <- unique(results_tera_sig$species)
tera_sig_spp <- tera_spp_abun %>%
  filter(species %in% tera_spp_list)
```

### Species that demonstrate a significant relationship with CTI from each experiment

```{r,message=F,echo=F,fig.width = 13, fig.height = 6}
# Plotting changes in abundance for significant species  
ggplot(tera_sig_spp, aes(x = mean_CTI, y = avg_abun, color = temp_treatment, group = temp_treatment)) +
  geom_smooth(method="lm") +
  geom_point() +
  theme_minimal() +
  labs(title = "TeRaCON: Mean CTI vs Abundance for Different Species",
       x = "Mean CTI",
       y = "Abundance") +
  scale_color_manual(values = c("HTamb" = "blue", "HTelv" = "red")) +
  facet_wrap(~ species, scales = "free_y")
ggplot(jrgce_sig_spp, aes(x = mean_CTI, y = avg_abun, color = temp_treatment, group = temp_treatment)) +
  geom_smooth(method="lm") +
  geom_point() +
  theme_minimal() +
  labs(title = "JRGCE: Mean CTI vs Abundance for Different Species",
       x = "Mean CTI",
       y = "Abundance") +
  scale_color_manual(values = c("ambient" = "blue", "warmed" = "red")) +
  facet_wrap(~ species, scales = "free_y")
ggplot(phace_sig_spp, aes(x = mean_CTI, y = avg_abun, color = temp_treatment, group = temp_treatment)) +
  geom_smooth(method="lm") +
  geom_point() +
  theme_minimal() +
  labs(title = "PHACE: Mean CTI vs Abundance for Different Species",
       x = "Mean CTI",
       y = "Abundance") +
  scale_color_manual(values = c("ambient" = "blue", "warmed" = "red")) +
  facet_wrap(~ species, scales = "free_y")
```
